### 2021-01-10 reports/results.Rmd
Updates supermarket results

---
title: "results"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "10/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::optlibrary(tidyverse)
knitr::opts_chunk$set(echo = FALSE)

df_results <- read.table("../run2-results-supermarket-sessions.csv", sep = ";",
                           header = TRUE)s_chunk$set(echo = TRUE)

df_ids <- read.table("../information.csv", sep = ";", header = TRUE)
df_rbans <- read.table("../information.csv", sep = ";", header = TRUE)
```