### 2021-01-10 reports/supermarket-output.Rmd

---
title: "Supermarket ouptut"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "10/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
df_results <- read.table("../run2-results-supermarket-sessions.csv", sep = ";",
                           header = TRUE)

df_demographics <- read.table("../demographics.csv", sep = ";", header = TRUE)
df_ids <- read.table("../information.csv", header = TRUE, sep = ";")

# Combine with demographics
df_join <- df_demographics %>%
  left_join(df_ids, by = "rbansid") %>%
  select(rbansid, vrid) %>% 
  filter(!is.na(rbansid) & !is.na(vrid))

df_results <- df_results %>%
  left_join(df_join, by = c("participant" = "vrid"))
```

```{r}
df_results <- df_results %>%
  mutate(total_picked_items = n_correct_items + n_extra_items,
         correct_ratio = n_correct_items/n_items,
         extra_ratio = n_extra_items/n_items) %>%
  group_by(participant, session) %>%
  mutate(n_trials = n()) %>%
  ungroup()

df_results <- filter(df_results, session <= 12)

df_results_remediation <- filter(df_results, type == "remediation")
df_results_test <- filter(df_results, type == "test")
```

```{r}
df_results_remediation %>%
  group_by(participant, session) %>%
  summarise(correct_ratio = mean(correct_ratio)) %>%
  ungroup() %>%
  ggplot(aes(session, correct_ratio)) +
    geom_line(aes(color=participant))
```
```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(correct_ratio = mean(correct_ratio), .groups = "drop_last") %>%
  summarise(correct_ratio = mean(correct_ratio)) %>%
  ggplot(aes(session, correct_ratio)) +
    geom_line() + scale_y_continuous(limits = c(0,1))
```

```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(n_trials = max(n_trials)) %>%
  ggplot(aes(session, n_trials)) +
    geom_point() + geom_smooth(method = "lm")
```

```{r}
df_results_remediation %>%
  group_by(session, participant) %>%
  summarise(n_trials = max(n_trials), .groups = "drop_last") %>%
  summarise(mean_trial = mean(n_trials)) %>%
  ggplot(aes(session, mean_trial)) +
    geom_line() + scale_y_continuous(limits = c(2,11))
```

```{r}
df_results_remediation %>%
  group_by(session, n_items) %>%
  summarise(results_time = mean(results_time)) %>%
  ggplot(aes(session, results_time)) +
  geom_line() + facet_wrap(~n_items)
```

```{r}
df_results_remediation %>%
  filter(results_time < 400) %>%
  ggplot(aes(session, results_time)) +
    geom_point() + geom_smooth(method = "lm") + 
  facet_wrap(~n_items)
```

```{r}
df_results_remediation %>%
  ggplot(aes(session, results_trajectory)) +
    geom_point() + geom_smooth(method = "lm") + 
  facet_wrap(~n_items)
```