### 2021-01-24 reports/rbans-output.Rmd
Updates existing rbans and supermarket output

---
title: "Rbans-output"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "15/10/2020"
output: html_document
---

```{r setup, include=FALSE}
library(brainvr.supermarket)
library(googlesheets4)
library(tidyverse)
library(knitr)
library(plotly)
knitr::opts_chunk$set(echo = FALSE)

source("../functions/cdz-loading-pilot.R")
source("../functions/cdz-analysis.R")
source("../functions/cdz-helpers.R")
source("../functions/cdz-getters.R")
source("../functions/cdz-preprocess.R")
```

```{r loading}
df_ids <- get_sheet_ids()
df_demographics <- get_sheet_demographics()
df_sessions <- get_sheet_sessions()
df_rbans <- get_sheet_rbans()
```

```{r preprocessing}
df_rbans <- add_session_to_rbans(df_rbans, df_demographics)

df_rbans_long <- df_rbans %>%
    pivot_longer(-c(rbansid, trenink, diagnosis, first_training, session),
                 names_to = "subtest", values_to = "score")

df_rbans_long <- df_rbans_long %>%
  group_by(trenink, first_training, subtest) %>%
  summarise(group_avg = mean(score, na.rm = TRUE)) %>%
  right_join(df_rbans_long, by = c("trenink", "first_training", "subtest"))
```
```{r}
df_ids %>%
  count(is_ok, rbans_ok, logy_ok) %>%
  kable(caption = "Overview of the state of affairs")

df_rbans_long %>%
  group_by(session) %>%
  summarise(length(unique(rbansid)))

df_ids %>%
  left_join(select(df_demographics, rbansid, first_training), by = "rbansid") %>%
  count(is_ok, first_training)
```

```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  ggplot(aes(subtest, score, fill = session)) +
  geom_boxplot() +
  geom_line(aes(x = session , y = group_avg, group=subtest), size = 1.25) +
  facet_wrap(~first_training)
```

```{r}
library(lme4)

df_rbans %>%
  lmer(iq_celek ~ session + first_training + (1|rbansid), data = .) %>%
  summary()

ggplot(df_rbans, aes(as.factor(session), iq_celek, fill=trenink)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  facet_wrap(~first_training)

df_rbans %>%
  filter(first_training == "TP") %>%
  glm(iq_celek ~ session ,data = .) %>%
  summary()

df_rbans %>%
  filter(first_training == "TP") %>%
  lmer(iq_celek ~ session + (1|rbansid), data = .) %>%
  summary()

df_rbans %>%
  filter(!is.na(first_training)) %>%
  ggplot(aes(session, vybaveniseznamu)) +
    geom_line(aes(color=rbansid)) + geom_point() +
    facet_wrap( ~ first_training + diagnosis)

df_rbans_baseline <- baseline_rbans(df_rbans)
df_rbans_baseline %>%
  pivot_longer(-c(rbansid, trenink, diagnosis, first_training), names_to = "subtest") %>%
  filter(grepl("iq_", subtest)) %>%
  ggplot(aes(subtest, value, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~first_training) + geom_hline(yintercept = 0)

```

```{r}
unique(df_rbans_long$subtest)
c("Vybavenifigury", "vybaveniseznamu", "vybavenipovidky")
```