### 2021-02-12 reports/questionnaire.Rmd
Updates questionnaire reports

---
title: "Questionnaire "
author: "Lukáš 'hejtmy' Hejtmánek"
date: "22/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
df_questionnaire <- read.table("../questionnaire.csv", sep = ";", header = TRUE) %>%
  mutate_at(vars(starts_with(c("vr", "paper", "supermarket"))), function(x){ifelse(x == 6, NA, x)})
df_demographics <- read.table("../demographics.csv", sep = ";", header = TRUE)

df_questionnaire <- df_questionnaire %>%
  left_join(df_demographics, c("rbansid" = "rbansid"))
```

```{r}
df_questionnaire_long <- df_questionnaire %>%
  pivot_longer(cols = c(starts_with(c("vr", "paper")), -ends_with("feedback")),
               names_to = "questionnaire_section", values_to = "score") %>%
  separate(questionnaire_section, sep = "_", into = c("question_focus", "question_area"),
           extra="merge")
```

```{r}
df_questionnaire_long %>%
  ggplot(aes(y = score, fill = question_focus)) + 
    geom_boxplot() + 
    facet_wrap(~question_area) +
    theme(legend.position="bottom")
```

```{r}
df_temp <- df_questionnaire_long %>%
  group_by(question_area, question_focus) %>%
  summarise(mean = mean(score, na.rm = TRUE)) %>%
  pivot_wider(names_from = question_focus, values_from = mean)

df_questionnaire_long %>%
  group_by(question_area) %>%
  summarise(t.test_p = t.test(score~question_focus)$p.value) %>%
  right_join(df_temp, by = "question_area") %>%
  knitr::kable(digits = 3)
```